How functional traits influence plant growth and shade tolerance across the life-cycle

### 2018-04-05 ms/figure_conceptual.tex
Tightening down to page limit
Revise paper after reviewers comments

\documentclass{beamer}
\usepackage[size=custom, width=21, height=28, scale=1.0, orientation=portrait]{beamerposter}
\setbeamertemplate{navigation symbols}{}
\setbeamersize{text margin left=0em,text margin right=0em}

@@

@@

\newcommand{\textbox}[2]{
  \node at (#1)[anchor=center] {
  {\footnotesize #2}};
}

@@

\node at ($(current page.north)+ (0cm, -7.5cm)$) (top){};

\node[rectangle, anchor = center] at ($(top)+ (0cm, 4cm)$){
  \includegraphics[width=18cm]{output/main/mass_fraction.pdf}};

\node at ($(top) + (-9.25cm, 6.75cm)$) [anchor=west, text width=12cm, text height=0.5cm]{
  {\bf \footnotesize Standing mass}};

\node at ($(top) + (-10.5cm, 6.75cm)$) [anchor=west, text width=12cm, text height=0.5cm]{
  {\bf \footnotesize a)}};

\node at ($(top) + (-10.5cm, 1.5cm)$) [anchor=west, text width=12cm, text height=0.5cm]{
  {\bf \footnotesize b)}};

@@

\defineRow{0}{thisrow}
\rowtitle{Biomass production}
\firstcol{1}{$\frac{{\rm d}B}{{\rm d}t}$}{output/conn/net_mass_production_dt_net_mass_production_dt}{}
\othercol{2}{\hspace{-1em} \scriptsize $\alpha_{\rm \tiny bio} \alpha_{\rm \tiny y}$}{output/conn/net_mass_production_dt_yield}{}
\othercol{3}{\hspace{-1em}$\times\Bigg($\hspace{-0.3em}\scriptsize$A_{\rm l} \bar{p}$}{output/conn/net_mass_production_dt_assimilation}{}
\othercol{4}{\hspace{-1em}-$\Sigma$\scriptsize $M_{\rm i} r_{\rm i}$}{output/conn/net_mass_production_dt_respiration}{\hspace{-0.5em}$\Bigg)$}
\othercol{5}{\hspace{-0.75em}-$\Sigma$\scriptsize $M_{\rm i} k_{\rm i}$}{output/conn/net_mass_production_dt_turnover}{}
\rowlabel{eq:dbdt_2}

@@

\textbox{$(top)+ (0cm, -20cm)$}{{\bf Plant height (m)}}

>>>>

\documentclass{beamer}
\usepackage[size=custom, width=21, height=29, scale=1.0, orientation=portrait]{beamerposter}
\setbeamertemplate{navigation symbols}{}
\setbeamersize{text margin left=0em,text margin right=0em}

@@

% Use square brackets for equations
\usepackage{mathtools}
\newtagform{brackets}{[}{]}
\usetagform{brackets}

@@

\newcommand{\textbox}[2]{
  \node at (#1)[anchor=center] {
  {\small #2}};
}

@@

\node at ($(current page.north)+ (0cm, -8.5cm)$) (top){};

\node[rectangle, anchor = center] at ($(top)+ (0.75cm, 5.25cm)$){
  \includegraphics[width=18cm]{output/main/mass_fraction.pdf}};

\node[rectangle, anchor = center, rotate=90, text width =6cm] at ($(top)+ (-8.75cm, 7cm)$){
  \small Fraction\\[-1ex] of mass
  };

\node at ($(top) + (-9.25cm, 7.75cm)$) [anchor=west, text width=12cm, text height=0.5cm]{
  {\bf \footnotesize Standing mass}};

\node at ($(top) + (-10.5cm, 7.75cm)$) [anchor=west, text width=12cm, text height=0.5cm]{
  {\bf \small a)}};

\node at ($(top) + (-10.5cm, 1.5cm)$) [anchor=west, text width=12cm, text height=0.5cm]{
  {\bf \small b)}};

@@

\defineRow{0}{thisrow}
\rowtitle{Biomass growth}
\firstcol{1}{$\frac{{\rm d}B}{{\rm d}t}$}{output/conn/net_mass_production_dt_net_mass_production_dt}{}
\othercol{2}{\hspace{-1em} \scriptsize $\,\alpha_{\rm \tiny bio} \alpha_{\rm \tiny y}$}{output/conn/net_mass_production_dt_yield}{}
\othercol{3}{\hspace{-1em}$\times\Bigg($\hspace{-0.3em}\scriptsize$A_{\rm l} \bar{p}$}{output/conn/net_mass_production_dt_assimilation}{}
\othercol{4}{\hspace{-1em}-$\Sigma$\scriptsize $M_{\rm i} r_{\rm i}$}{output/conn/net_mass_production_dt_respiration}{\hspace{-0.5em}$\Bigg)$}
\othercol{5}{\hspace{-0.75em}-$\Sigma$\scriptsize $M_{\rm i} k_{\rm i}$}{output/conn/net_mass_production_dt_turnover}{}
\rowlabel{eq:dbdt_2}

@@

\textbox{$(top)+ (0cm, -20cm)$}{Plant height (m)}
\textbox{$(top)+ (0cm, 2.5cm)$}{Plant height (m)}

### 2018-04-05 ms/ms-common-defs.tex
Revise paper after reviewers comments

\usepackage[rgb,dvipsnames]{xcolor}
\definecolor{grey}{rgb}{0.5, 0.5, 0.5}

\newcommand{\smurl}[1]{\url{#1}}
\newcommand{\ud}{\ensuremath{\rm{d}}}
\newcommand{\tabitem}{~~\llap{\textbullet}~~}
\newcommand{\email}[1]{\href{mailto:#1}{\texttt{#1}}}

\title{Untangling the link between traits, size and growth rate in plants}
\runninghead{I\lowercase{nfluence of traits on growth}}

\author{Daniel S. Falster\textasteriskcentered \& Richard G. FitzJohn}

\affiliation{Biological Sciences, Macquarie University NSW 2109, Australia
\\
\textasteriskcentered Correspondence author. E-mail: \texttt{adaptive.plant@gmail.com}
}

\usepackage{natbib}
\bibliographystyle{ecol_let}
\setcitestyle{authoryear,open={(},close={)}}
% allows for numbered referencing, as required by ecol letters
\usepackage{etoolbox}
\newbool{MyRefNumbers}
\booltrue{MyRefNumbers} % comment to remove numbers in reference list

@@

% We will generate all images so they have a width \maxwidth. This means
% that they will get their normal width if they fit onto the page, but
% are scaled down if they would overflow the margins.
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
\else\Gin@nat@width\fi}
\makeatother

\let\Oldincludegraphics\includegraphics
\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}

>>>>

% abbreviations
\newcommand{\smurl}[1]{\url{#1}}
\newcommand{\ud}{\ensuremath{\rm{d}}}
\newcommand{\tabitem}{~~\llap{\textbullet}~~}
\newcommand{\email}[1]{\href{mailto:#1}{\texttt{#1}}}
\newcommand{\plant}{\texttt{plant}}
\newcommand{\wplcp}{\textsc{wplcp}}
\newcommand{\lma}{\textsc{lma}}
\newcommand{\wood}{\textsc{wd}}
\newcommand{\seed}{\textsc{sm}}
\newcommand{\hmat}{\textsc{h}$_{\rm mat}$}
\newcommand{\nitrogen}{\textsc{n}}
\newcommand{\sma}{\textsc{sma}}

@@

% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
\author[a,1]{Daniel S. Falster}
\author[b]{Remko A. Duursma}
\author[a,c]{Richard G. FitzJohn}

% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
\affil[a]{Evolution \& Ecology Research Centre, and School of Biological, Earth and
Environmental Sciences, University of New South Wales, Sydney NSW 2052, Australia}
\affil[b]{Hawkesbury Institute for the Environment, Western Sydney University, Locked Bag 1797, Penrith NSW 2751, Australia}
\affil[c]{Imperial College, London, United Kingdom}

% Please give the surname of the lead author for the running footer
\leadauthor{Falster}

### 2018-04-05 ms/ms-supp.Rnw
Add longtable package
minor edits, address change
Revise paper after reviewers comments

\documentclass[10pt,twoside]{article}

\usepackage{suppmat}
\usepackage{listings}
\usepackage[T1]{fontenc} % Better underscores

@@

% We will generate all images so they have a width \maxwidth. This means
% that they will get their normal width if they fit onto the page, but
% are scaled down if they would overflow the margins.
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}

\titleprefix{Supporting Materials for:}

% Figure command -- also adds figure to table of contents
\newcommand{\myfigure}[4]{
    \begin{figure}
            \centering
            \includegraphics{#1}
            \caption{\textbf{#2} #3}
            \label{#4}
    \end{figure}
  \addcontentsline{toc}{subsection}{Fig \ref{#4} -- #2}
}

\date{}

<<preload, echo=FALSE>>=
library(knitr)
library(xtable)
opts_chunk$set(echo=FALSE)
@

@@

\tableofcontents

\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\thetable}{S\arabic{table}}
\setcounter{secnumdepth}{0}

\clearpage

\section{A -- Additional details on the \texttt{FF16} growth model in the {\plant} package}

In this paper we use the \texttt{FF16} growth model from the package {\plant} \citep{Falster-2016}, which is accessed via \textsc{r} \citep{R-2015}. A full derivation of  Eqns 1-21 from the main text are described is provided in Appendix of \citet{Falster-2016}, see also \citep{Falster-2011}. Below we repeat some additional details from \citet{Falster-2016}, so that readers have a complete picture of the model.

@@

{\plant} allows for the hyper-parameterisation of the \texttt{FF16} physiological model via plant functional traits: this enables simultaneous variation in multiple parameters in accordance with an assumed trade-off. The hyper-parameterisation function used in the current analysis differs slightly from that in \citet{Falster-2016}. The functions implemented are as follows.

The trait leaf mass per unit area, denoted by $\phi$, directly influences growth by changing ${\rm d}A_{\rm l} / {\rm d}M_{\rm a}$. In addition, we link $\phi$ to the rate of leaf turnover, based on a widely observed scaling relationship from \citet{Wright-2004},
$$k_{\rm l} = \beta_{{\rm kl1}} \, \left(\frac{\phi}{\phi_0}\right)^{-\beta_{{\rm kl2}}}.$$
This relationship is normalised around $\phi_0$, the global mean of $\phi$. This allows us to vary $\beta_{{\rm kl1}}$ and $\beta_{{\rm kl2}}$ without displacing the relationship from the observed mean.

We also vary the mass-based leaf respiration rate so that it stays constant per unit leaf area and varies with $\phi$ and nitrogen per unit leaf area $\nu$, as empirically observed by \citet{Wright-2004},
$$r_{\rm l} = \frac{\beta_{{\rm lf4}}\, \nu}{\phi}.$$

The trait wood density, denoted by $\rho$, directly influences growth by changing ${\rm d}A_{\rm l} / {\rm d}M_{\rm a}$. In addition, we allow for $\rho$ to influence the rate of growth-independent mortality,
$$d_{\rm I} = \beta_{{\rm dI1}} \, \left(\frac{\rho}{\rho_0}\right) ^ {-\beta_{{\rm dI2}}},$$
and also the rate of sapwood turnover,
$$k_{\rm s} = \beta_{{\rm ks1}} \, \left(\frac{\rho}{\rho_0}\right)^ {-\beta_{{\rm ks2}}}.$$
As for $\phi$, these relationships are normalized around $\rho_0$, the global mean of $\rho$. By default, $\beta_{{\rm kI2}}$ and $\beta_{{\rm ks2}}$ are set to zero, so these linkages only become present when these parameters are set to something other than their default values.

The rate of sapwood respiration per unit volume is assumed to be constant, so sapwood respiration per unit mass varies as
$$r_{\rm s}  = \frac{\beta_{{\rm rs1}}}{\rho},$$
where $\beta_{{\rm rs1}}$ is a default rate per volume of sapwood. Similarly, the rate of bark respiration per unit mass varies as
$$r_{\rm b} = \frac{\beta_{{\rm rb1}}}{\rho},$$
with $\beta_{{\rm rb1}} = 2 \beta_{{\rm rs1}}$.

@@

Effects of the trait seed mass, denoted by $\omega$, are naturally embedded in the equation determining fecundity (Eq. \ref{eq:fecundity}) and the initial height of seedlings. In addition, we let the accessory cost per seed be a multiple of seed size,
$$\alpha_{{\rm f3}} = \beta_{{\rm f1}} \omega,$$
as empirically observed \citep{Henery-2001}.

@@

Photosynthesis per unit leaf area and respiration rates per unit leaf mass (or area) are assumed to vary with leaf nitrogen per unit area, $\nu$. The calculation of respiration rates is already described above. To calculate the average annual photosynthesis for a leaf, we integrate the instantaneous rate per unit leaf area over the annual solar trajectory, using a rectangular-hyperbolic photosynthesis light response curve,

$$p(\nu, E) = \frac1{365 \textrm{d}}\int_0^{365 \textrm{d}}\frac{Y(t) + A_{\rm max} - \sqrt{(Y(t) + A_{\rm max})^2 - 4 \beta_{{\rm lf2}} Y(t) A_{\rm max} }}{2 \beta_{{\rm lf2}}} {\rm d}t,$$
where
\begin{itemize}
\item $A_{\rm max}$ is the maximum photosynthetic capacity of the leaf,
\item $\beta_{{\rm lf2}}$ is the curvature of the light response curve,
\item $Y(t) = \beta_{{\rm lf3}} I(t)$ is the initial yield of the light response curve, with $\beta_{{\rm lf3}}$ being the quantum yield parameter,
\item $I(t) = k_{\rm I} \, I_0(t)\, E$ is the intensity of light on the leaf surface, and
\item $I_0(t)$ is light incident on a surface perpendicular to the sun's rays directly above the canopy at time $t$.
\end{itemize}
The profile of $I_0(t)$ is given by a solar model adapted from \citet{TerSteege-1997}.

@@

@@

<<echo=FALSE>>=
make_hyperpar2
@

\clearpage

\section{B -- Supporting Tables}\label{app:supp_info_table}

\addcontentsline{toc}{subsection}{Table \ref{tab:params_core} -- Core parameters of the \texttt{FF16} physiological model}

<<echo=FALSE,results='asis'>>=
tab <- table_plant_parameters("../data/growth_model_pars_core.csv")
print(xtable(tab,
            hline.after=c(1),
            align='lp{7cm}llll',
            label="tab:params_core",
            caption="Core parameter of the \texttt{FF16} physiological model."),
      sanitize.text.function=I,
      include.rownames=FALSE, caption.placement = "top")
@

@@

\addcontentsline{toc}{subsection}{Table \ref{tab:params_hyper} -- Parameters for hyper-parameterisation of the \texttt{FF16} physiological model}

<<echo=FALSE,results='asis'>>=
tab <- table_plant_parameters("../data/growth_model_pars_hyper.csv")
print(xtable(tab,
            hline.after=c(1),
            align='lp{7cm}llll',
            label="tab:params_hyper",
            caption="Parameters for hyper-parameterisation of the \\texttt{FF16} physiological model."),
      sanitize.text.function=I,
      include.rownames=FALSE, caption.placement = "top")
@

\clearpage

\section{C -- Supporting Figures}\label{app:supp_info_figures}

@@

@@

\clearpage

\addcontentsline{toc}{section}{References}
\bibliography{references}

>>>>

\documentclass[9pt,onecolumn,twoside]{pnas-new}
% Use the lineno option to display guide line numbers if required.
% Note that the use of elements such as single-column equations
% may affect the guide line number alignment.

\templatetype{pnasresearcharticle}

\title{
{\normalfont \fontsize{12pt}{16pt}
{\color{black50} Supporting information for:}\\} \vskip6pt
\fontsize{16pt}{20pt} How functional traits influence plant growth and shade tolerance across the life-cycle
}

@@

\dates{This supplement was compiled on \today}
\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}

\hypersetup{
     linkcolor=pnasbluetext
}

@@

\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\thetable}{S\arabic{table}}
\renewcommand{\theequation}{S\arabic{equation}}
\renewcommand{\thesection}{S\arabic{section}}

% Figure command -- also adds figure to table of contents
\newcommand{\myfigure}[4]{
    \begin{figure*}[!hb]
            \centering
             \includegraphics[width=0.9\textwidth]{#1}
            \caption{\textbf{#2} #3}
            \label{#4}
    \end{figure*}
}

\section{Additional details on the \texttt{FF16} growth model in the {\plant} package}

In this paper we use the \texttt{FF16} growth model from the package {\plant} \citep{Falster-2016}, which is accessed via \textsc{r} \citep{R-2015}. A full derivation of  Eqns 1-21 from the main text are described is provided in Appendix of \citep{Falster-2016}, see also \citep{Falster-2011}. Below we repeat some additional details from \citep{Falster-2016}, so that readers have a complete picture of the model.

@@

{\plant} allows for the hyper-parameterisation of the \texttt{FF16} physiological model via plant functional traits: this enables simultaneous variation in multiple parameters in accordance with an assumed trade-off. The hyper-parameterisation function used in the current analysis differs slightly from that in \citep{Falster-2016}. The functions implemented are as follows.

@@

Effects of the trait seed mass, are naturally embedded in the equation determining fecundity and the initial height of seedlings. In addition, we let the accessory cost per seed be a multiple of seed size,
$\alpha_{{\rm f3}} = \beta_{{\rm f1}} \omega,$
as empirically observed \citep{Henery-2001}.

@@

Photosynthesis per unit leaf area and respiration rates per unit leaf mass (or area) are assumed to vary with leaf nitrogen per unit area, $\nu$. The calculation of respiration rates is already described above. To calculate the average annual photosynthesis for a leaf, we integrate the instantaneous rate per unit leaf area over the annual solar trajectory, using a rectangular-hyperbolic photosynthesis light response curve,
\begin{equation} \label{eq:pts}
p(\nu, E) = \frac1{365 \textrm{d}}\int_0^{365 \textrm{d}}\frac{Y(t) + A_{\rm max} - \sqrt{(Y(t) + A_{\rm max})^2 - 4 \beta_{{\rm lf2}} Y(t) A_{\rm max} }}{2 \beta_{{\rm lf2}}} {\rm d}t,
\end{equation}
where $A_{\rm max}$ is the maximum photosynthetic capacity of the leaf, 
$\beta_{{\rm lf2}}$ is the curvature of the light response curve,
$Y(t) = \beta_{{\rm lf3}} I(t)$ is the initial yield of the light response curve, with $\beta_{{\rm lf3}}$ being the quantum yield parameter,
$I(t) = k_{\rm I} \, I_0(t)\, E$ is the intensity of light on the leaf surface, and
$I_0(t)$ is light incident on a surface perpendicular to the sun's rays directly above the canopy at time $t$.
The profile of $I_0(t)$ is given by a solar model adapted from \citep{TerSteege-1997}.

@@

The trait leaf mass per unit area, denoted by $\phi$, directly influences growth by changing ${\rm d}A_{\rm l} / {\rm d}M_{\rm a}$. In addition, we link $\phi$ to the rate of leaf turnover, based on a widely observed scaling relationship from \citep{Wright-2004},
\begin{equation}
k_{\rm l} = \beta_{{\rm kl1}} \, \left(\frac{\phi}{\phi_0}\right)^{-\beta_{{\rm kl2}}}.
\end{equation}
This relationship is normalised around $\phi_0$, the global mean of $\phi$. This allows us to vary $\beta_{{\rm kl1}}$ and $\beta_{{\rm kl2}}$ without displacing the relationship from the observed mean.

We also vary the mass-based leaf respiration rate so that it stays constant per unit leaf area and varies with $\phi$ and nitrogen per unit leaf area $\nu$, as empirically observed by \citep{Wright-2004},
\begin{equation}
r_{\rm l} = \frac{\beta_{{\rm lf4}}\, \nu}{\phi}.
\end{equation}

The trait wood density, denoted by $\rho$, directly influences growth by changing ${\rm d}A_{\rm l} / {\rm d}M_{\rm a}$. As for $\phi$, these relationships are normalized around $\rho_0$, the global mean of $\rho$. By default, $\beta_{{\rm kI2}}$ and $\beta_{{\rm ks2}}$ are set to zero, so these linkages only become present when these parameters are set to something other than their default values.

The rate of sapwood respiration per unit volume is assumed to be constant, so sapwood respiration per unit mass varies as
\begin{equation}
r_{\rm s}  = \frac{\beta_{{\rm rs1}}}{\rho},
\end{equation}
where $\beta_{{\rm rs1}}$ is a default rate per volume of sapwood. Similarly, the rate of bark respiration per unit mass varies as
\begin{equation}
r_{\rm b} = \frac{\beta_{{\rm rb1}}}{\rho},
\end{equation}
with $\beta_{{\rm rb1}} = 2 \beta_{{\rm rs1}}$.

@@

<<echo=FALSE, highlight=TRUE, comment = ''>>=
make_hyperpar2
#txt <- capture.output(make_hyperpar2)
#writeLines(strwrap(txt, width = 150, exdent = 10))
@

\section{References}
\bibliography{references}

\section{Supporting Tables and Figures}\label{app:supp_info_table}

<<echo=FALSE,results='asis'>>=
tab <- table_plant_parameters("../data/growth_model_pars_core.csv")
print(xtable(tab,
            hline.after=c(1),
            align='lp{7cm}llll',
            label="tab:params_core",
            caption="Core parameter of the \texttt{FF16} physiological model."),
      floating.environment="table*",
      sanitize.text.function=I,
      include.rownames=FALSE, 
      caption.placement = "top")
@

<<echo=FALSE,results='asis'>>=
tab <- table_plant_parameters("../data/growth_model_pars_hyper.csv")
print(xtable(tab,
            hline.after=c(1),
            align='lp{9cm}llll',
            label="tab:params_hyper",
            caption="Parameters for hyper-parameterisation of the \\texttt{FF16} physiological model."),
      sanitize.text.function=I,
      floating.environment="table*",
      include.rownames=FALSE, 
      caption.placement = "top")
@

@@

@@

\myfigure{main/assumptions.pdf}{Assumptions of a functional balance model for plant construction.}{Each dot is a single plant from the \textsc{baad} \citep{Falster-2015b}.
Blue lines show standardised major axis lines fit to different species. The black line shows the relationship assumed in this paper, with slope given by the functional-balance assumptions in Table \ref{tab:allometry}.}{fig:assumptions}

@@
